Detecting Hate Speech with GPT-3

### 2021-03-23 outputs/documents/paper.tex
Fix one typo and save tex file

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: true
    toc: false
    fig_caption: true
    latex_engine: xelatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
  -  \usepackage{amsmath}
biblio-style: apalike
title: "Detecting Hate Speech with GPT-3"
thanks: "Code and data are available at: https://github.com/kelichiu/GPT3-hate-speech-detection. We gratefully acknowledge the support of Gillian Hadfield and the Schwartz Reisman Institute for Technology and Society. We thank Amy Farrow, Haoluan Chen, Mauricio Vargas Sepúlveda, and Tom Davidson for helpful suggestions. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto and Schwartz Reisman Institute
abstract: "Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist? We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 is able to identify sexist or racist text with an accuracy between 48 per cent and 69 per cent. With few-shot learning and an instruction included in the prompt, the model's accuracy can be as high as 78 per cent. We conclude that large language models have a role to play in hate speech detection, and that with further development language models could be used to counter hate speech and even self-police."
keywords: "GPT-3; natural language processing; quantitative analysis; hate speech."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
endnote: no
graphics: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), echo = FALSE, warning = FALSE)
# library(dplyr)
library(tidyverse)
# library(tidyr)
# library(readr)
```

# Introduction

Sophisticated natural language processing (NLP) models, such as OpenAI's GPT-3, can produce hateful text. In particular, there have been many examples of text being generated that targets marginalized groups based on their sex, race, sexual orientation, and other characteristics. Large language models are trained on enormous datasets from various sources. This means that untruthful statements, human biases, and abusive language are inevitably included. Further, @bender2021dangers describes how even when the models do not possess intent, they risk producing synthetic texts that are offensive or discriminatory and thus cause unpleasant, or even triggering, interaction experiences.

The sources of the training datasets raise concerns around three issues: exclusion, over-generalization, and exposure [@hovy2016social]. Exclusion happens due to the demographic bias in the dataset. In the case of language models that are trained on US and UK English scraped from the Internet, datasets may be disproportionately white, male, and young. Therefore, it is not surprising to see white supremacist, misogynistic, and ageist content being over-represented in training datasets [@bender2021dangers]. Over-generalization stems from the assumption that what we see in the dataset represents what occurs in the real world. Words such as 'always', 'never', 'everybody', or 'nobody' are frequently used for rhetorical purpose instead of their literal meanings. However, language models do not always recognize this and make inference based on generalized statements using these words. For instance, hate speech commonly uses generalized language for targeting a group such as 'all' and 'every', and a model trained on these statements may generate similarly generalized and harmful statements. Finally, exposure refers to the relative attention, and hence considerations of importance, given to something. In the context of NLP this may be reflected in the emphasis on English-language created under particular circumstances, rather than another languages or circumstances that may be more prevalent.

While these, and other, issues give us pause, the dual-use problem, which is that the same technology can be applied to both good and bad uses, provides motivation. For instance, while stylometric analysis can reveal the identity of political dissenters, it can also solve the unknown authorship of historic text [@hovy2016social]. In this paper we are interested in whether given large language models can produce harmful language, can they also identify, or learn to identify, harmful language?

Even though large NLP models do not have a real understanding of language, the vocabularies and the construction patterns of harmful languages can be thought of as known to them. We show that this knowledge can be used to identify abusive language and even hate speech. In particular we consider 120 different extracts that have been categorized as 'racist', 'sexist', or 'neither' in the single-category settings and 240 different extracts in the mixed-category settings. We ask GPT-3 to classify these based on zero-, one-, and few-shot learning, with and without instruction. We find that the model performs best with few-shot learning when an instruction is included. In that setting the model is able to accurately classify around 78 per cent of the extracts. If language models can be used to identify abusive language, then not only is there potential for them to counter the production of abusive language by humans, but they could also potentially self-police.

# Background

## Language models, Transformers and GPT-3

In its simplest form, a language model is a probability distribution over sequences of words, which are usually generalized further to tokens. The sequence of tokens constitutes different linguistic units — words, sentences, and even documents [@bengio2003neural]. Language models predict the next token based on inputs. If we consider each token in a vocabulary as a dimension, then the dimensionality of language quickly becomes large [@rosenfeld2000two]. Over time a variety of statistical language models have been created to nonetheless enable prediction. The n-gram is one of the earliest language models. It works by considering the co-occurrence of tokens in a sequence. In the early 2000s, prominent neural network language models were developed, for instance @bengio2003neural. These were then built on by word embeddings language models in the 2010s in which the distance between tokens represents how related those tokens are, for instance @turian2010word. In 2017, @vaswani2017attention introduced the Transformer, which marked a new era for language models. The Transformer is a network architecture for neural networks that can be trained more quickly than many other approaches [@vaswani2017attention]. Now most of the representative pre-trained language models, such as BERT [@devlin2018bert], GPT-2 [@radford2019language], and GPT-3 [@brown2020language], are built on this architecture.

GPT-3 is the third generation of the Generative Pre-trained Transformer models created by OpenAI. Until January 11, when @fedus2021switch announced a Transformer model with a trillion parameters, GPT-3 was the largest, publicly-known, Transformer language model. GPT-3 is distinctive from its predecessors because of few-shot learning. This means that GPT-3 can 'learn' to perform a new task based on only a few examples, expressed in natural language, instead of a fine-tuning process that can require a large amount of data. GPT-3 has led to unexpected NLP applications, such as computational code generation given natural language prompts. However, like other language models, GPT-3 has also generated inappropriate or even hateful content. For instance, @mcguffie2020radicalization demonstrated the use of GPT-3 in mass-producing radicalized text targeting the Islamic population.

## Hate speech detection

Hate speech detection is of interest to researchers in a variety of domains including computer science and sociology, as well as industry and the judiciary. Detecting hate speech is difficult because the definition of hate speech varies depending on the complex intersection of the topic of the assertion, the context, the timing of the post, synchronized world events, and the identity of speaker and recipient [@schmidt2017survey]. Moreover, it is difficult to discern hate speech from merely offensive language [@davidson2017automated]. Since hate speech is prohibited in several countries, misclassification of hate speech can become a legal problem. For instance, in Canada, speech that contains 'public incitement of hatred' or 'wilful promotion of hatred' is specified by the Criminal Code [@act2021justice]. Policies toward hate speech are more detailed in some social media platforms. For instance, the Twitter Hateful Conduct Policy states:

> You may not promote violence against or directly attack or threaten other people on the basis of race, ethnicity, national origin, caste, sexual orientation, gender, gender identity, religious affiliation, age, disability, or serious disease. We also do not allow accounts whose primary purpose is inciting harm towards others on the basis of these categories.
>
> @twitterpolicy2017

There has been a large amount of research focused on detecting hate speech. And as part of this various hate speech datasets have been created and examined. For instance, @waseem2016hateful create a dataset that captures hate speech in the form of racist and sexist language that includes domain expert annotation, and @davidson2017automated trains a classifier to distinguish between hate speech and offensive language. And it is important to note that even these datasets have bias. For instance, @davidson2019racial found racial bias in five different sets of Twitter data annotated for hate speech and abusive language. They found that tweets written in African American English are more likely to be labeled as abusive.

# Methods

@@

## Dataset

We use the ETHOS dataset created by @mollas2020ethos. ETHOS is based on comments found in YouTube and Reddit. The ETHOS YouTube data is collected through Hatebusters [@anagnostou2018hatebusters]. Hatebusters is a platform that collects comments from YouTube and assigns a 'hate' score to them using a support vector machine. The Reddit data is collected from the Public Reddit Data Repository [@baumgartner2020pushshift]. The dataset has two variants: binary and multi-label. In the binary dataset comments are classified as hate or non-hate. In the multi-label variant, the comments are evaluated on measures that include violence, gender, race, ability, religion, and sexual orientation. The examples provided in this paper are from the ETHOS dataset and hence contain typos, misspelling, and offensive content.

@@

## Zero-shot learning

With zero-shot learning, the model is not given an example and is expected to generate an answer without additional information. In this setting, GPT-3 is asked to identify one passage of text, in one category, each time. A simple instruction is given asking the model to answer only 'Yes' or 'No'. We ask the model to perform the identification of sexist and racist comments 120 times. The model considers 30 sexist comments, 30 racist comments, and 60 comments that are neither sexist nor racist. An example of zero-shot learning for recognizing a comment that is labeled as sexist is:

@@

> Is the following text sexist? Answer yes or no.
>
> 'The thing is women are not equal to us men and their place is the home and kitchen'
>
> Answer:

@@

> Yes.

## One-shot learning

With one-shot learning, similarly to the zero-shot learning case, the model is asked to identify one passage of text in one category at a time. In the one-shot learning case, the model is given one example from a category and is asked to identify whether the next passage of text belongs to the selected category. Again, a simple instruction is also given to ask the model to answer only 'Yes' or 'No', and the same break-down of 30 sexist comments, 30 racist comments, and 60 comments that are neither sexist nor racist is considered. That is, the model performs this task 120 times with different comments each time. An example of one-shot learning for recognizing a comment that is labeled as sexist is:

@@

> The following text in quotes is sexist:
> 
> 'Feminism is a very terrible disease'
> 
> Is the following text sexist? Answer yes or no.
> 
> 'She is heavily relying on him to turn the other cheek...tough talking demon infested woman.'
> 
> Answer:

@@

> Yes.

## Few-shot learning

Under the few-shot setting, three examples are provided to the model. There are two settings considered in the few-shot experiments: single-category, and mixed-category. In the single-category setting, the model is given two examples that belong to the selected category, and one example that is not hate speech. The answer is expected to be binary. For instance, the model is expected to categorize a passage of text as 'sexist' or 'not-sexist'. An example of this is:

@@

> 'Too bad women don’t know how to kill themselves': sexist.
>
> 'You should use your time to arrest murderers not little kids': not-sexist.
> 
> 'Now they know better than this shit lol they dudes. The stronger sex. The man supremacy': sexist.
> 
> 'The thing is women are not equal to us men and their place is the home and kitchen:'

@@

> sexist.

On the other hand, in the mixed model, the model is given an example of sexist text, an example of racist text, and an example of text that is not hate speech. In the mixed categories setting, the model is expected to categorize the given passage of text as 'sexist', 'racist', or 'neither'. There are ten sets of examples; each set of examples is used for the model to classify 60 racist comments, 60 sexist comments, and 120 comments that are in neither category This results in 2,400 responses in total. The ten sets of examples are available in Table \@ref(tab:appendixa) in the Appendix. An example of this is:

@@

> 'That girl still need a good ass whooping, stupid ass bitch!!': sexist. 
>
> 'U SUCK SHIT YOU BLACK FUCK': racist. 
>
> 'You deserve way more subscribers. The shit you say is fucking gold.': neither. 
>
> 'Every girl on this planet curses her mother for their birth. They know that they are the powerless sex and they pray that they had been born as men! With dicks and stuff.':

@@

> sexist.

Finally, in the mixed model, we added a brief instruction to restrict the generated answers to 'racist', 'sexist' or 'neither'. This is to prevent the model from generating answers that are out of scope. An example of this is:

@@

> Classify the following texts into 'racist', 'sexist' or 'neither'. 
>
> 'That girl still need a good ass whooping, stupid ass bitch!!': sexist. 
>
> 'U SUCK SHIT YOU BLACK FUCK': racist. 
>
> 'You deserve way more subscribers. The shit you say is fucking gold.': neither. 
>
> 'Every girl on this planet curses her mother for their birth. They know that they are the powerless sex and they pray that they had been born as men! With dicks and stuff.':

@@

> sexist.

# Result

## Zero-shot learning

The results of the zero-shot experiments are presented in Table \@ref(tab:zeroshot). The model has 35 matches and 25 mismatches in the sexist category, and 23 matches and 37 mismatches in the racist category. The model performs better when identifying sexist comments compared with identifying racist comments. However, the overall ratio of matches and mismatches is 58:62. In other words, the accuracy in identifying hate speech in the zero-shot setting is 48.3 per cent.

```{r zeroshot}
zero_shot_result <- read.csv(here::here("outputs/data/zero_shot_results.csv"))

zero_shot_result <- 
  zero_shot_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups="drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

zero_shot_result %>% 
  knitr::kable(caption = "Classification of statements with zero-shot learning",
               booktabs = TRUE)
```

## One-shot learning

The results of the one-shot learning experiments are presented in Table \@ref(tab:oneshot). The model has 46 matches and 14 mismatches in the racist category, and 37 matches and 23 mismatches in the sexist category. In contrast with the result generated from zero-shot learning, the model performs slightly better in identifying racist comments compared with identifying sexist comments. The general performance in the one-shot setting is also better than in the zero-shot setting. The overall ratio of matches and mismatches is 83:37. In other words, the accuracy of identifying hate speech in the one-shot setting is 69.2 per cent.

```{r oneshot}
one_shot_result <- read.csv(here::here("outputs/data/one_shot_results.csv"))

one_shot_result <- 
  one_shot_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups = "drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

one_shot_result %>% 
  knitr::kable(caption = "Classification of statements with one-shot learning", 
             booktabs = TRUE)
```

## Few-shot learning -- single category

>>>>

\documentclass[12pt,]{article}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}

  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother

\setcounter{secnumdepth}{0}

\usepackage{longtable,booktabs}

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

\title{Detecting Hate Speech with GPT-3 \thanks{Code and data are available at: \url{https://github.com/kelichiu/GPT3-hate-speech-detection}. We gratefully acknowledge the support of Gillian Hadfield and the Schwartz Reisman Institute for Technology and Society. We thank Amy Farrow, Haoluan Chen, Mauricio Vargas Sepúlveda, and Tom Davidson for helpful suggestions. Comments on the 22 March 2021 version of this paper are welcome at: \href{mailto:rohan.alexander@utoronto.ca}{\nolinkurl{rohan.alexander@utoronto.ca}}.}  }

\author{\Large Ke-Li Chiu\vspace{0.05in} \newline\normalsize\emph{University of Toronto}   \and \Large Rohan Alexander\vspace{0.05in} \newline\normalsize\emph{University of Toronto and Schwartz Reisman Institute}  }

\date{}

\usepackage{titlesec}

\titleformat*{\section}{\normalsize\bfseries}
\titleformat*{\subsection}{\normalsize\itshape}
\titleformat*{\subsubsection}{\normalsize\itshape}
\titleformat*{\paragraph}{\normalsize\itshape}
\titleformat*{\subparagraph}{\normalsize\itshape}

\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage[strings]{underscore} % protect underscores in most circumstances

\newtheorem{hypothesis}{Hypothesis}
\usepackage{setspace}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

% move the hyperref stuff down here, after header-includes, to allow for - \usepackage{hyperref}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Ke-Li Chiu (University of Toronto) and Rohan Alexander (University of Toronto and Schwartz Reisman Institute)},
             pdfkeywords = {GPT-3; natural language processing; quantitative analysis; hate speech.},  
            pdftitle={Detecting Hate Speech with GPT-3},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----

% add tightlist ----------
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}

\begin{document}

% \pagenumbering{arabic}% resets `page` counter to 1 
%

% \maketitle

{% \usefont{T1}{pnc}{m}{n}
\setlength{\parindent}{0pt}
\thispagestyle{plain}
{\fontsize{18}{20}\selectfont\raggedright 
\maketitle  % title \par

}

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont Ke-Li Chiu} \hskip 15pt \emph{\small University of Toronto}   \par \textbf{\authorfont Rohan Alexander} \hskip 15pt \emph{\small University of Toronto and Schwartz Reisman Institute}

}

}

\begin{abstract}

    \hbox{\vrule height .2pt width 39.14pc}

    \vskip 8.5pt % \small

\noindent Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist? We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 is able to identify sexist or racist text with an accuracy between 48 per cent and 69 per cent. With few-shot learning and an instruction included in the prompt, the model's accuracy can be as high as 78 per cent. We conclude that large language models have a role to play in hate speech detection, and that with further development language models could be used to counter hate speech and even self-police.

\vskip 8.5pt \noindent \emph{Keywords}: GPT-3; natural language processing; quantitative analysis; hate speech. \par

    \hbox{\vrule height .2pt width 39.14pc}

\end{abstract}

\vskip -8.5pt

 % removetitleabstract

\noindent

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Sophisticated natural language processing (NLP) models, such as OpenAI's GPT-3, can produce hateful text. In particular, there have been many examples of text being generated that targets marginalized groups based on their sex, race, sexual orientation, and other characteristics. Large language models are trained on enormous datasets from various sources. This means that untruthful statements, human biases, and abusive language are inevitably included. Further, \citet{bender2021dangers} describes how even when the models do not possess intent, they risk producing synthetic texts that are offensive or discriminatory and thus cause unpleasant, or even triggering, interaction experiences.

The sources of the training datasets raise concerns around three issues: exclusion, over-generalization, and exposure \citep{hovy2016social}. Exclusion happens due to the demographic bias in the dataset. In the case of language models that are trained on US and UK English scraped from the Internet, datasets may be disproportionately white, male, and young. Therefore, it is not surprising to see white supremacist, misogynistic, and ageist content being over-represented in training datasets \citep{bender2021dangers}. Over-generalization stems from the assumption that what we see in the dataset represents what occurs in the real world. Words such as `always', `never', `everybody', or `nobody' are frequently used for rhetorical purpose instead of their literal meanings. However, language models do not always recognize this and make inference based on generalized statements using these words. For instance, hate speech commonly uses generalized language for targeting a group such as `all' and `every', and a model trained on these statements may generate similarly generalized and harmful statements. Finally, exposure refers to the relative attention, and hence considerations of importance, given to something. In the context of NLP this may be reflected in the emphasis on English-language created under particular circumstances, rather than another languages or circumstances that may be more prevalent.

While these, and other, issues give us pause, the dual-use problem, which is that the same technology can be applied to both good and bad uses, provides motivation. For instance, while stylometric analysis can reveal the identity of political dissenters, it can also solve the unknown authorship of historic text \citep{hovy2016social}. In this paper we are interested in whether given large language models can produce harmful language, can they also identify, or learn to identify, harmful language?

Even though large NLP models do not have a real understanding of language, the vocabularies and the construction patterns of harmful languages can be thought of as known to them. We show that this knowledge can be used to identify abusive language and even hate speech. In particular we consider 120 different extracts that have been categorized as `racist', `sexist', or `neither' in the single-category settings and 240 different extracts in the mixed-category settings. We ask GPT-3 to classify these based on zero-, one-, and few-shot learning, with and without instruction. We find that the model performs best with few-shot learning when an instruction is included. In that setting the model is able to accurately classify around 78 per cent of the extracts. If language models can be used to identify abusive language, then not only is there potential for them to counter the production of abusive language by humans, but they could also potentially self-police.

\hypertarget{background}{%
\section{Background}\label{background}}

\hypertarget{language-models-transformers-and-gpt-3}{%
\subsection{Language models, Transformers and GPT-3}\label{language-models-transformers-and-gpt-3}}

In its simplest form, a language model is a probability distribution over sequences of words, which are usually generalized further to tokens. The sequence of tokens constitutes different linguistic units --- words, sentences, and even documents \citep{bengio2003neural}. Language models predict the next token based on inputs. If we consider each token in a vocabulary as a dimension, then the dimensionality of language quickly becomes large \citep{rosenfeld2000two}. Over time a variety of statistical language models have been created to nonetheless enable prediction. The n-gram is one of the earliest language models. It works by considering the co-occurrence of tokens in a sequence. In the early 2000s, prominent neural network language models were developed, for instance \citet{bengio2003neural}. These were then built on by word embeddings language models in the 2010s in which the distance between tokens represents how related those tokens are, for instance \citet{turian2010word}. In 2017, \citet{vaswani2017attention} introduced the Transformer, which marked a new era for language models. The Transformer is a network architecture for neural networks that can be trained more quickly than many other approaches \citep{vaswani2017attention}. Now most of the representative pre-trained language models, such as BERT \citep{devlin2018bert}, GPT-2 \citep{radford2019language}, and GPT-3 \citep{brown2020language}, are built on this architecture.

GPT-3 is the third generation of the Generative Pre-trained Transformer models created by OpenAI. Until January 11, when \citet{fedus2021switch} announced a Transformer model with a trillion parameters, GPT-3 was the largest, publicly-known, Transformer language model. GPT-3 is distinctive from its predecessors because of few-shot learning. This means that GPT-3 can `learn' to perform a new task based on only a few examples, expressed in natural language, instead of a fine-tuning process that can require a large amount of data. GPT-3 has led to unexpected NLP applications, such as computational code generation given natural language prompts. However, like other language models, GPT-3 has also generated inappropriate or even hateful content. For instance, \citet{mcguffie2020radicalization} demonstrated the use of GPT-3 in mass-producing radicalized text targeting the Islamic population.

\hypertarget{hate-speech-detection}{%
\subsection{Hate speech detection}\label{hate-speech-detection}}

Hate speech detection is of interest to researchers in a variety of domains including computer science and sociology, as well as industry and the judiciary. Detecting hate speech is difficult because the definition of hate speech varies depending on the complex intersection of the topic of the assertion, the context, the timing of the post, synchronized world events, and the identity of speaker and recipient \citep{schmidt2017survey}. Moreover, it is difficult to discern hate speech from merely offensive language \citep{davidson2017automated}. Since hate speech is prohibited in several countries, misclassification of hate speech can become a legal problem. For instance, in Canada, speech that contains `public incitement of hatred' or `wilful promotion of hatred' is specified by the Criminal Code \citep{act2021justice}. Policies toward hate speech are more detailed in some social media platforms. For instance, the Twitter Hateful Conduct Policy states:

\begin{quote}
You may not promote violence against or directly attack or threaten other people on the basis of race, ethnicity, national origin, caste, sexual orientation, gender, gender identity, religious affiliation, age, disability, or serious disease. We also do not allow accounts whose primary purpose is inciting harm towards others on the basis of these categories.

\citet{twitterpolicy2017}
\end{quote}

There has been a large amount of research focused on detecting hate speech. And as part of this various hate speech datasets have been created and examined. For instance, \citet{waseem2016hateful} create a dataset that captures hate speech in the form of racist and sexist language that includes domain expert annotation, and \citet{davidson2017automated} trains a classifier to distinguish between hate speech and offensive language. And it is important to note that even these datasets have bias. For instance, \citet{davidson2019racial} found racial bias in five different sets of Twitter data annotated for hate speech and abusive language. They found that tweets written in African American English are more likely to be labeled as abusive.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

@@

\hypertarget{dataset}{%
\subsection{Dataset}\label{dataset}}

We use the ETHOS dataset created by \citet{mollas2020ethos}. ETHOS is based on comments found in YouTube and Reddit. The ETHOS YouTube data is collected through Hatebusters \citep{anagnostou2018hatebusters}. Hatebusters is a platform that collects comments from YouTube and assigns a `hate' score to them using a support vector machine. The Reddit data is collected from the Public Reddit Data Repository \citep{baumgartner2020pushshift}. The dataset has two variants: binary and multi-label. In the binary dataset comments are classified as hate or non-hate. In the multi-label variant, the comments are evaluated on measures that include violence, gender, race, ability, religion, and sexual orientation. The examples provided in this paper are from the ETHOS dataset and hence contain typos, misspelling, and offensive content.

@@

\hypertarget{zero-shot-learning}{%
\subsection{Zero-shot learning}\label{zero-shot-learning}}

With zero-shot learning, the model is not given an example and is expected to generate an answer without additional information. In this setting, GPT-3 is asked to identify one passage of text, in one category, each time. A simple instruction is given asking the model to answer only `Yes' or `No'. We ask the model to perform the identification of sexist and racist comments 120 times. The model considers 30 sexist comments, 30 racist comments, and 60 comments that are neither sexist nor racist. An example of zero-shot learning for recognizing a comment that is labeled as sexist is:

@@

\begin{quote}
Is the following text sexist? Answer yes or no.

`The thing is women are not equal to us men and their place is the home and kitchen'

Answer:
\end{quote}

@@

\begin{quote}
Yes.
\end{quote}

\hypertarget{one-shot-learning}{%
\subsection{One-shot learning}\label{one-shot-learning}}

With one-shot learning, similarly to the zero-shot learning case, the model is asked to identify one passage of text in one category at a time. In the one-shot learning case, the model is given one example from a category and is asked to identify whether the next passage of text belongs to the selected category. Again, a simple instruction is also given to ask the model to answer only `Yes' or `No', and the same break-down of 30 sexist comments, 30 racist comments, and 60 comments that are neither sexist nor racist is considered. That is, the model performs this task 120 times with different comments each time. An example of one-shot learning for recognizing a comment that is labeled as sexist is:

@@

\begin{quote}
The following text in quotes is sexist:

`Feminism is a very terrible disease'

Is the following text sexist? Answer yes or no.

`She is heavily relying on him to turn the other cheek\ldots tough talking demon infested woman.'

Answer:
\end{quote}

@@

\begin{quote}
Yes.
\end{quote}

\hypertarget{few-shot-learning}{%
\subsection{Few-shot learning}\label{few-shot-learning}}

Under the few-shot setting, three examples are provided to the model. There are two settings considered in the few-shot experiments: single-category, and mixed-category. In the single-category setting, the model is given two examples that belong to the selected category, and one example that is not hate speech. The answer is expected to be binary. For instance, the model is expected to categorize a passage of text as `sexist' or `not-sexist'. An example of this is:

@@

\begin{quote}
`Too bad women don't know how to kill themselves': sexist.

`You should use your time to arrest murderers not little kids': not-sexist.

`Now they know better than this shit lol they dudes. The stronger sex. The man supremacy': sexist.

`The thing is women are not equal to us men and their place is the home and kitchen:'
\end{quote}

@@

\begin{quote}
sexist.
\end{quote}

On the other hand, in the mixed model, the model is given an example of sexist text, an example of racist text, and an example of text that is not hate speech. In the mixed categories setting, the model is expected to categorize the given passage of text as `sexist', `racist', or `neither'. There are ten sets of examples; each set of examples is used for the model to classify 60 racist comments, 60 sexist comments, and 120 comments that are in neither category This results in 2,400 responses in total. The ten sets of examples are available in Table \ref{tab:appendixa} in the Appendix. An example of this is:

@@

\begin{quote}
`That girl still need a good ass whooping, stupid ass bitch!!': sexist.

`U SUCK SHIT YOU BLACK FUCK': racist.

`You deserve way more subscribers. The shit you say is fucking gold.': neither.

`Every girl on this planet curses her mother for their birth. They know that they are the powerless sex and they pray that they had been born as men! With dicks and stuff.':
\end{quote}

@@

\begin{quote}
sexist.
\end{quote}

Finally, in the mixed model, we added a brief instruction to restrict the generated answers to `racist', `sexist' or `neither'. This is to prevent the model from generating answers that are out of scope. An example of this is:

@@

\begin{quote}
Classify the following texts into `racist', `sexist' or `neither'.

`That girl still need a good ass whooping, stupid ass bitch!!': sexist.

`U SUCK SHIT YOU BLACK FUCK': racist.

`You deserve way more subscribers. The shit you say is fucking gold.': neither.

`Every girl on this planet curses her mother for their birth. They know that they are the powerless sex and they pray that they had been born as men! With dicks and stuff.':
\end{quote}

@@

\begin{quote}
sexist.
\end{quote}

\hypertarget{result}{%
\section{Result}\label{result}}

\hypertarget{zero-shot-learning-1}{%
\subsection{Zero-shot learning}\label{zero-shot-learning-1}}

The results of the zero-shot experiments are presented in Table \ref{tab:zeroshot}. The model has 35 matches and 25 mismatches in the sexist category, and 23 matches and 37 mismatches in the racist category. The model performs better when identifying sexist comments compared with identifying racist comments. However, the overall ratio of matches and mismatches is 58:62. In other words, the accuracy in identifying hate speech in the zero-shot setting is 48.3 per cent.

\begin{table}

\caption{\label{tab:zeroshot}Classification of statements with zero-shot learning}
\centering
\begin{tabular}[t]{llr}
\toprule
Result & Category & Count\\
\midrule
Match & Racist & 23\\
Match & Sexist & 35\\
Mismatch & Racist & 37\\
Mismatch & Sexist & 25\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{one-shot-learning-1}{%
\subsection{One-shot learning}\label{one-shot-learning-1}}

The results of the one-shot learning experiments are presented in Table \ref{tab:oneshot}. The model has 46 matches and 14 mismatches in the racist category, and 37 matches and 23 mismatches in the sexist category. In contrast with the result generated from zero-shot learning, the model performs slightly better in identifying racist comments compared with identifying sexist comments. The general performance in the one-shot setting is also better than in the zero-shot setting. The overall ratio of matches and mismatches is 83:37. In other words, the accuracy of identifying hate speech in the one-shot setting is 69.2 per cent.

\begin{table}

\caption{\label{tab:oneshot}Classification of statements with one-shot learning}
\centering
\begin{tabular}[t]{llr}
\toprule
Result & Category & Count\\
\midrule
Match & Racist & 46\\
Match & Sexist & 37\\
Mismatch & Racist & 14\\
Mismatch & Sexist & 23\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{few-shot-learning-single-category}{%
\subsection{Few-shot learning -- single category}\label{few-shot-learning-single-category}}

The results of the single-category, few-shot learning, experiments are presented in Table \ref{tab:fewshotsingle}. The model has 41 matches and 19 mismatches in the racist category, and 42 matches and 18 mismatches in the sexist category. Here, the model performs almost equally well at identifying racist comments compared with identifying sexist comments. The general performance in the few-shot learning setting is similar to the performance in the one-shot learning setting. The overall ratio of matches and mismatches is 83:37, or around 69.2 per cent, however the composition is different, compared with the one-shot learning setting.

\begin{table}

\caption{\label{tab:fewshotsingle}Classification of statements with single-category few-shot learning}
\centering
\begin{tabular}[t]{llr}
\toprule
Result & Category & Count\\
\midrule
Match & Racist & 41\\
Match & Sexist & 42\\
Mismatch & Racist & 19\\
Mismatch & Sexist & 18\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{few-shot-learning-mixed-category}{%
\subsection{Few-shot learning -- mixed category}\label{few-shot-learning-mixed-category}}
